---
title: "Thesis Project Proposal Meeting"
author: "Maggie Slein"
date: "9/13/20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goals for today's meeting 
1. Discuss project proposal below and the different ways to go
2. 
3. Outline timeline for regular meetings and when milestones for this project should accomplished
4. Discuss appropriate workflow for this project 


# Main Objectives: 
What is the single main objective of the research? This needs to relate to some kind of general ecological issue, and be articulated in a single concise statement.

How is environmental variation (temperature, pH, nutrient composition) systematically studied and manipulated across organismal, population, and community ecology and how does study design influence conclusions about environmental variation compare across these bodies of literature?

Does environmental variation affect how rates scale across levels of biological organization?




# Testable hypothesis: A hypothesis or set of hypotheses must:
clearly contribute to achieving the main objective
be able to be definitively tested and rejected
ideally, the set of hypotheses should together support some inference about the main objective.
Generating hypotheses is a challenge. Make sure there are not logical gaps between the
hypotheses, objective and methods used to test them, or if gaps exist, address them.

H0: There is no difference in how variation is studied at different levels of biological organization and thus the conclusions are the same across all levels.

Ha: Short term responses to simple environmental variation manipulation is prioritized at the organismal level, long term and more complex environmental variation manipulated is prioritized at the population and community level, owing to differential experimental design and conclusions about how variation type will interact in the ecosystem.

H02: Environmental variation does affect how rate scale across levels of biological organization?



# Methods:

1. I will conduct a series of ISI literature searches to get a hollistic set of studies referencing environmental variation with selected keywords


How will hypotheses be tested? General categories include experiments, observations (surveys),
meta-analyses, analysis of theoretical models. Will inferences be drawn from more than one type
of evidence?
- How will data be collected and analyzed? What statistical tests and software will you use?
- What resources are required? Budget for time (e.g., timeline) and for costs.
#Results: Consider, in advance, the 2-3 key figures or tables you will produce for your final
paper. Sketch them, including axes, and even the data that would reject or fail your hypotheses or
support your predictions.
#Inference and Next Steps: What logical steps will you use to achieve your main objectives,
and draw conclusions? How exactly will your figures (sketched above) answer your original
questions? Consider different possible versions of the results, and how you might interpret them.
Then revisit your questions – are you happy with how your hypotheses are stated, given the
possible results?
```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
