---
title: 'My Final College Paper'
author: 'Margaret A. Slein'
date: 'May 2021'
institution: 'Reed College'
division: 'Mathematics and Natural Sciences'
advisor: 'Sam Fey'
department: 'Biology'
degree: 'Bachelor of Arts'
knit: bookdown::render_book
site: bookdown::bookdown_site
params:
  'Install needed packages for {thesisdown}': True
output:
  thesisdown::thesis_pdf: default 
abstract: '`r if(knitr:::is_latex_output()) paste(readLines(here::here("prelims", "00-abstract.Rmd")), collapse = "\n  ")`'
acknowledgements: |
  
  The future is bright, the future is dark 
 
  Time stands still, yet flies like a lark 
 
  Memories endure and memories fade 
 
  People are rough, yet smooth like suede
 
  The days seem long, the days seem short 
 
  It rains all year, until the sunshine retorts 
 
  Reed is an experience, Reed is a place 
 
  Reed is a journey of learning how to keep pace
 
  Reed is a individual, Reed is a kollective 
 
  Reed is space for being reflective 
 
  Reed is the present, Reed is the past
  
  Reed is the net for our future that we cast


  To those from Reed 
 
  To those from before 
 
  To those from the outside
 
  To those at the core 
 
  To those I still talk to 
 
  To those I talk to no more
 
  To those still here today
 
  To those who now soar 
 
  A few words of dear thanks….


  Thank you for the laughs, the tears, the screams 
 
  Gratitude for the late nights, the dancing, the dreams 
 
  Thank you for the walks, the talks, and meals together 
 
  Many thanks for the shenanigans I’ll cherish forever 
 
  Thank you for the compassion, for the coffee and tea 
 
  Appreciation for the love, the hate, and learning to be 

  Thank you for the support, and most ardently, for embracing me 

dedication: |
  You can have a dedication here if you wish. 
bibliography: bib/thesis.bib
csl: csl/apa.csl
lot: true
lof: true
---

<!--
Above is the YAML (YAML Ain't Markup Language) header that includes a lot of 
metadata used to produce the document.  Be careful with spacing in this header!

If you'd prefer to not include a Dedication, for example, simply delete the section entirely, or silence them (add # before each line). 

If you have other LaTeX packages you would like to include, delete the # before header-includes and list the packages after hyphens on new lines.

If you'd like to include a comment that won't be produced in your resulting file enclose it in a block like this.

If you receive a duplicate label error after knitting, make sure to delete the index.Rmd file and then knit again.
-->

```{r include_packages, include=FALSE}
# This chunk ensures that the thesisdown package is
# installed and loaded. This thesisdown package includes
# the template files for the thesis.
if (!require(remotes)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("remotes", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste('You need to run install.packages("remotes")",
            "first in the Console.')
    )
  }
}
if (!require(thesisdown)) {
  if (params$`Install needed packages for {thesisdown}`) {
    remotes::install_github("ismayc/thesisdown")
  } else {
    stop(
      paste(
        "You need to run",
        'remotes::install_github("ismayc/thesisdown")',
        "first in the Console."
      )
    )
  }
}
library(thesisdown)
# Set how wide the R output will go
options(width = 70)
```

<!--
The acknowledgments, preface, dedication, and abstract are added into the PDF
version automatically by inputting them in the YAML at the top of this file.
Alternatively, you can put that content in files like 00--prelim.Rmd and
00-abstract.Rmd like done below.
-->



```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00--prelim.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00-abstract.Rmd")}

```

<!-- The {.unnumbered} option here means that the introduction will be 
"Chapter 0." You can also use {-} for no numbers on chapters.
-->

# Introduction {.unnumbered}

## The historical significance of variability 
It is no secret that the natural world has relied on resilience and persistence in the face of environmental change for millenia (cite). Cyanobacteria can trace their origin back to a bacteria that evolved the ability to perform oxygenic photosynthesis to persist in a changing world (Soo et al 2017, science). Similarly, vascular plants have evolved different photosynthetic systems (C3, CAM) to manage a variety of thermal and spatial conditions (cite). While cyanobacteria and vascular plants evolved new strategies and structures to persist in the environment, the environment has continued to remain dynamic, such that evolutionary responses and plasticity may become even more relevant (cite). Understanding how organisms respond to the environment and its dynamic nature is significant for how to capitulate ecosystems currently and in the future.

## Relevant significance of variability today 
In addition to the importance of environmental variability across geologic timescales, organisms also interact with environmental variability on several shorter timescales. Ecosystems and organisms inhabiting those environments have experienced variation on many scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns (i.e. ENSO). The scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems (cite). As environmental patterns continue to change, from increased mean temperatures to more severe weather events (i.e. hurricanes), understanding the limits to organisms’ ability to cope with changes in environmental conditions is increasingly important (cite). 

## Why this study on variability matters 
While there is a large body of literature on environmental variability including its relevance for X, y, and Z (include citation), there is a lack of consensus as to which scales of biological organization will be most severely affected by different shifts in environmental conditions (problem statement). We aim to better understand how environmental variability affects the magnitude of responses across organization levels as well as whether different patterns of variation differentially affect different levels of organization (aims). To accomplish this, we review recent progress in the field of environmental variability, specifically changes in temperature, to identify gaps between theory, experiment, and implementation. 

